\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-18 standardmodel.tex
Adapt style guide.

Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory. We prove the consistency by extending the standard model from \citep{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1. we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2. we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma. We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e. the judgementally-equal syntax needs to be mapped to the same data for well-defined-ness).

@@

We encourage the reader to read the (fake) Agda-style proof in our appendix because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation. Here, we need to clarify some ambiguity in our conventional formulation. For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on $\denotesS{\Sigma}$'', the best way to understand is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ also defined inductively on the length of the signature. The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$. \textbf{This is actually the formulation in Agda} -- QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement type also interpreted.

Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data -- it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc. What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-defined-ness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.

Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together -- because $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.

@@

We sketch the model for the inductive type and inheritance judgement here, and both of them require inductive facility in meta-logic. In set theory, that would be using \textit{Rule sets}~\citep{timany2017consistency,aczel1998relating}, but we are using Agda's Inductive facility instead. Both of them are somewhat ``interpreted by themselves'' in Agda, because their syntactic definition is pretty ``well-founded''. For inductive type, for each grounded signature, we have an inductive type using introduction rule of W-type as constructor, shown in Figure \ref*{fig:model-ind-type}. This "MWsup" when applying to $n$ will use the $n$-th constructor specify in the signature. This grounded-ness is required to verify the substitution laws. For inheritance judgement, we simply construct a term model of inheritance judgement using vanilla inductive type (instead of QIIT). Then we use meta-level substitution to model the explicit substitution, and we can define $\denotesS{"inh"~\_ ~\_}$ by induction on this term model.

@@

The verification of validity of this model is largely omitted here and we refer the reader to the appendix. To verify our model preserve the judgemental equality, usually unpacking the definition is enough -- for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$ by function extensionality. One exception is for $\cC$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together. The compatibility lemmas can be verified directly -- for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{p_1\mu~o}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.

>>>>

Once we have the syntax of our theory, the first question is if our theory is consistent, i.e., if we can syntactically derive bottom in our theory. We prove the consistency by extending the standard model from~\cite{altkap2016,kaposi2017type, kaposi2019gluing} and their formalization, where they formally resort to the concept of \textit{algebra} of QIIT, here we only consider inductively interpreting each syntax piece into naive set theory that is also respecting judgemental equalities\footnote{In our appendix, we still interpret into Agda component to be more precise and less error-prone, also we need the inductive facility from Agda.} The main idea of the proof is that: 1. we use $\denotesS{}$ to map each intrinsic well-typed syntax piece into a mathematical object; 2. we define proof-irrelevant proposition $\models_S$ for each syntactical $\vdash$, and derive compatibility lemma. We will mainly focus on the definition of $\denotesS{}$ and omit the verification of $\models_S$\footnote{Actually our 1 is acting exactly as an interpretation function, and in our agda formulation we only do 1 because 2 is automatic; here we make 2 explicit to make our formulation more accessible and uniform}, and also we omit the verification of equality (i.e. the judgementally-equal syntax needs to be mapped to the same data for well-definedness).

@@

We encourage the reader to read the (fake) Agda-style proof in our appendix because most of the parts are similar but with more type annotation so it is clearer about the mathematical object doing the interpretation. Here, we need to clarify some ambiguity in our conventional formulation. For example, when we wrote ``$\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ defined inductively on $\denotesS{\Sigma}$'', the best way to understand is to consider we are defining a function $\denotesS{\goodType{\Gamma}{\cC \_}{}}$ as the interpretation for $\goodType{\Gamma}{\cC \_}{}$ that mapping $\denotesS{\sigma}$ to $\denotesS{\goodType{\Gamma}{\cC \sigma}{}}$ also defined inductively on the length of the signature. The case is similar for $\denotesS{\cL \_}$, $\denotesS{\cCt \_}$ and  $\denotesS{\Sigma \ \_ \ \_}$. \textbf{This is actually the formulation in Agda}---QIIT syntax consider $\cC : "Sig"~\Gamma~n \to "Ty"~\Gamma$ as a constructor and our model needs to interpret this constructor to a function $\cC^S : "Sig"^S \Gamma^S~n \to "Ty"^S~\Gamma^S$ with each judgement type also interpreted.

Basically in Agda, we are simply constructing a function that mapping our QIIT data-type to some other data---it maps $"Con"$ to $"Con"^S$, "Ty : Con → Set" to "Ty$^S$ : Con$^S$ → Set" and etc. What's more, since judgemental equality in our meta-theory is represented by equality in QIIT, and the well-definedness of this mapping-out function implies that ``judgementally equal syntax is interpreted by the same semantic in the model''.

Notice that $\denotesS{\goodSig{\Gamma}{s}{n}}$ and $\denotesS{\goodType{\Gamma}{\cC\sigma}{i}}$ is mutual recursively defined together---because $\goodSealS{\Gamma}{f}{\sigma}{A}$ relies on $\cC$.

@@

We sketch the model for the inductive type and inheritance judgement here, and both of them require inductive facility in meta-logic. In set theory, that would be using \textit{Rule sets}~\cite{timany2017consistency,aczel1998relating}, but we are using Agda's Inductive facility instead. Both of them are somewhat ``interpreted by themselves'' in Agda, because their syntactic definition is pretty ``well-founded''. For inductive type, for each grounded signature, we have an inductive type using introduction rule of W-type as constructor, shown in \cref{fig:model-ind-type}. This "MWsup" when applying to $n$ will use the $n$-th constructor specify in the signature. This grounded-ness is required to verify the substitution laws. For inheritance judgement, we simply construct a term model of inheritance judgement using vanilla inductive type (instead of QIIT). Then we use meta-level substitution to model the explicit substitution, and we can define $\denotesS{"inh"~\_ ~\_}$ by induction on this term model.

@@

The verification of validity of this model is largely omitted here and we refer the reader to the appendix. To verify our model preserve the judgemental equality, usually unpacking the definition is enough---for example, to verify $\beta$-rule for linkage projection, we only need to see that $\denotesS{\goodTerm{\Gamma}{p_1\mu~(\mu^+~o~t)}{\cL\sigma}} = \denotesS{\goodTerm{\Gamma}{o}{\cL\sigma}}$ by function extensionality. One exception is for $\cC$ and $\goodSig{\_}{\_}{n}$, the substitution law around these two are also proved together by induction because they were mutual-recursively defined together. The compatibility lemmas can be verified directly---for example $\frac{\goodSigS{\Gamma}{\sigma}{n+1}\quad\goodTermS{\Gamma}{o}{\cL\sigma}}{\goodTermS{\Gamma}{p_1\mu~o}{\cL(p_1\nu~\sigma)}}$ is proved by simply unfolding the definition.

### 2022-06-18 syntactic-example.tex
Save.
Adapt style guide.

We provide one example in its surface syntax as pseudocode, and their meta-theory encoding. This example is computing the predecessor of a mundane natural number -- however we split the inductive definition into two parts and making "O" and "S" standing alone. "pred" is defined as $n \mapsto ("pred" n, n)$, and thus a predecessor.

\begin{figure}[H]\label{fig:example-pseudocode}
\begin{minipage}[t]{0.5\linewidth}
\begin{minted}{Coq}
Family A.
 Inductive N := O : N. 
 (* Osig, sig₁, obj₁, sig₂,
    O₁, obj₂, sig₂', sl₁ *)
 Family handler.
   O :self[A].N * self[A].N 
     := (self[A].O, self[A].O).

 EndFamily. (* pN₁, handler₁, sig₃ *)
 pred 
  : self[A].N → self[A].N×self[A].N 
  := ... 
  (* predT, pred₁, sig₄, sig₄', sl₄ *) 
 k := self[A].pred self[A].O 
(* pN₂, sig₅ *)
EndFamily.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.45\linewidth}
\begin{minted}{Coq}
Family A₂ extends A.
 Extend Inductive N := S : N -> N.
 (* Nsig, sig₁₁, inh₁, O₂, sig₂₁, inh₂, 
    sig₂₂, sig₂₁', sig₂₂', sl₂₂, inh₃ *)
 Extend Family handler.
   Inherit O. (* inhᵢₙ₀, pN₃ *)
   S : N*N -> N*N 
     := fun (x,y) => (y, self[A₂].S y).
 EndFamily. (* inhᵢₙ, sig₃₁, inh₄*)
 pred 
  : self[A₂].N → self[A₂].N×self[A₂].N 
  := ... 
  (* sig₄₁, inh₅, sl₄₁ *)
 Inherits k. 
 (* inh₆ *)
EndFamily.
    \end{minted}
  \end{minipage}
  \caption{Demonstrating Pseudo-Code}
\end{figure}

@@

Here we explain how  our detailed construction in Figure \ref{fig:example-construction} is constructed. We first elaborate that how our inductive type "N" is written in the specific format (the style of W-type). We use 1, 0 here to indicate unit type and bottom type.

>>>>

We provide one example in its surface syntax as pseudocode, and their meta-theory encoding. This example is computing the predecessor of a mundane natural number, but we split the inductive definition into two parts and making "O" and "S" standing alone. "pred" is defined as $n \mapsto ("pred" n, n)$, and thus a predecessor.

\begin{figure}[H]\label{fig:example-pseudocode}
\begin{minipage}[t]{0.5\linewidth}
\begin{minted}{Coq}
Family A.
 FInductive N := O : N. 
 (* Osig, sig₁, obj₁, sig₂,
    O₁, obj₂, sig₂', sl₁ *)
 Family handler.
   O :self[A].N × self[A].N 
     := (self[A].O, self[A].O).

 EndFamily. (* pN₁, handler₁, sig₃ *)
 pred 
  : self[A].N → self[A].N × self[A].N 
  := ... 
  (* predT, pred₁, sig₄, sig₄', sl₄ *) 
 k := self[A].pred self[A].O 
(* pN₂, sig₅ *)
EndFamily.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.45\linewidth}
\begin{minted}{Coq}
Family A₂ extends A.
 Extend FInductive N := S : N → N.
 (* Nsig, sig₁₁, inh₁, O₂, sig₂₁, inh₂, 
    sig₂₂, sig₂₁', sig₂₂', sl₂₂, inh₃ *)
 Extend Family handler.
   Inherit O. (* inhᵢₙ₀, pN₃ *)
   S : self[A].N × self[A].N 
       → self[A].N × self[A].N 
     := fun (x,y) => (y, self[A₂].S y).
 EndFamily. (* inhᵢₙ, sig₃₁, inh₄*)
 pred 
  : self[A₂].N → self[A₂].N×self[A₂].N 
  := ... 
  (* sig₄₁, inh₅, sl₄₁ *)
 Inherits k. 
 (* inh₆ *)
EndFamily.
    \end{minted}
  \end{minipage}
  \caption{Demonstrating Pseudo-Code}
\end{figure}

@@

Here we explain how  our detailed construction in Figure \cref{fig:example-construction} is constructed. We first elaborate that how our inductive type "N" is written in the specific format (the style of W-type). We use 1, 0 here to indicate unit type and bottom type.

### 2022-06-18 syntax.tex
1. Recursor rules are added into metatheory 2. Coq Implementation section is done 3. The two Example is basically done, I only need to add statistics. Please Check.
Adapt style guide.

Firstly, there are four standard judgement forms in MLTT,
well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
$\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. Our presentation
follows the style of \citep{kaposi2019gluing}, but for readability we
omit universe levels attached to judgements.
After these,
we will have four new kinds of judgement to model our linkage.

@@

Due to our intrinsic setting, \textbf{we can and will omit a lot of presumptions of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly -- because it is not possible to have "not-well-formed" context in our setting.

@@

\judgebox{\goodWSig{\Gamma}{\sigma}{n}}
$$
\Rule[name=Emp WSig]
{}
{\goodWSig{\Gamma}{w\cdot}{0}}
\quad
\Rule[name=WSig Add]
{\goodWSig{\Gamma}{s}{n}
  \quad \goodType{\Gamma}{A}{i}
  \quad \goodType{\Gamma, A}{B}{i}}
{\goodWSig{\Gamma}{w^+ \  s \  A \  B}{n+1}}
\quad
\Rule[name=Ind Univ]
{\goodWSig{\Gamma}{\sigma}{n}}
{\goodType{\Gamma}{\bW \sigma}{i}}
$$
%\YZ{Is there an introduction rule for terms of type 'bW sigma'?}
%\EDJreply{Yes there is. I omit it because it is a singleton type, please see my newly added explanation below.}

@@

Recall in W-type \citep{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type. To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}. Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type, and given an inductive type signature, we use \ruleref{WSig Proj} to extract and corresponding $A_i, B_i$.

Of course we need to put an eye on the substitution law for each piece of the syntax. Here $\bW \sigma$ is a large singleton type \cite{stone2000}, and it only has one ``element'', the inductive type itself, inside. We formulate in this cumbersome way because we want to have inductive type to be a field element and exposing the constructors at the type level for later \textit{pattern-matching coverage checking} (i.e. if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed).\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type. They are basically the same thing but (W T) cannot be at the type position in a judgement, i.e. k |- q : (W T) is not allowed, we only have k |- q : El (W T). This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}\YZreply{I think explanation is warranted; readers unfamiliar with that line of work may find it confusing. BTW, is the elimination rule for "El (W T)" omitted?}\EDJreply{Please check the following paragraph for the explanation for El, and also Univ-1 is the elimination rule for "El T" in general. I just didn't write out the two equations c (El T) = T, El (c T) = T}. We follow the style of \textit{type universe à la Tarski} and distinguish types from their \textit{codes or names}, and thus we use $El\ T$ as the type given type name $T : \cU$ -- in these cases, $T$ is not allowed to locate at type position (i.e. at the right-most position of a term judgement) but $El \ T$ is.

@@

The signature (for the linkage) and the linkage rules are quite similar the signature for inductive type.  They are both constructed from the empty stuff, and respect substitution lemma for sure. We have \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each componenet due to harmony \citep{pfenning2009lecture} thus we have local soundness and local completeness ($\beta,\eta$ rules). We omit them and the substitution laws for either field addition and projection. For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible\footnote{Interestingly, this omission can be supported by Agda using Implicit Arguments}.

@@

In our meta-theory, we only formulate non-dependent eliminator, i.e. the
\textit{recursor}. Each recursor is constructed by a linkage containing
functions/handlers/pattern-matching cases dealing with each
constructor. Thus code reuse for recursor is delegated to linkage
inheritance---to reuse one particular pattern-matching case,
we just inherit the function/handler. This greatly simplifies the
metatheory development.

@@

Now, one missing piece for linkage is the sealing judgement, and both of the $\mu^+, \nu^+$ rules relies on that -- then what is this sealing judgement for?

@@

As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
a short hand of term judgement $\goodTerm{\Gamma , \cC
\sigma}{\_}{A[\pi_1]}$. The substitution law should be understood as a
lemma, and it is easily derivable using $\gamma^\uparrow$.

The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance -- due to the presence of inductive type in the context, not everything is inheritable. For example, $ X_1 : \bW \sigma_1 \vdash t : T$ in the parent family will generally not be able to be inherited to the children once we override the inductive type. In other words, there is no way ``transplanting'' $t$ into the context $X_2 : \bW \sigma_2$. This makes all the fields after this inductive type non-inheritable -- which shouldn't be the case because a lot of fields might not be directly relying on this concrete $\bW \sigma_1$.

Sealing is used to exploit this fact -- with sealing, we can add well-typed terms $X_1 : \cU \vdash t : T$, into the parent family, in the context of $ X_1 : \bW \sigma_1$. Later when we want to inherit $t$, because $t$ only consider $X_1$ as an arbitrary type, inheritance is plausible.

Of course, any fields using "Wsup" and "Wrec" directly relies on the $\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and \ruleref{Rec-Constr}) and thus non-inheritable -- in these cases, "Wsup" can be understood as concrete constructors and "Wrec" are \textit{exhaustiveness checking}. These fields together with inductive types are the one sealing responsible for hiding. After compilation of linkage type into existential type, we can easily hide the \textit{concrete} definition of the inductive type while exposing all the fields using "Wsup" and "Wrec" like an abstract interface. Apparently, future fields relies on this abstract interface will have no problem on inheritance. \YZ{Does coverage checking mean exhaustiveness checking?}\EDJreply{Yes. The one make sure every case is handled in pattern-matching. Do you want me to replace the terminology using exhaustiveness checking? }\YZreply{Yes, I think exhaustiveness is standard parlance.}
Of course, since "Wsup" and "Wrec" directly rely on the concrete inductive definition, we will have to override these fields whenever inductive type is overridden. For "Wsup" we only need to override each constructor and for "Wrec" it is just doing exhaustiveness checking inside each children family. The recursor handlers however, can be safely inherited.
\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
ML, sealing has the connotation of hiding things behind an existential
type. So using the same word may cause the reader to have false
presumptions about what is going on here.}\EDJreply{You are right. Our "Sealing" is not hiding a particular field into an existential type, but our seal is make several fields together into an "existential type". Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}\EDJreply{I will resolve this comment once I add the throughout example and you find it clear -- because I haven't changed any text}

@@

With the definition of sealing judgement in mind, we can see
\ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the
implementation. In the implementation, we will have "self__" all around
when defining new fields in our plugin, and this "self__" is exactly the
$A$ in the context of the two fields -- especially when we don't do any
sealing, this $A = \cC \sigma$, which is exactly how the implementation
exposes "self__" as a module parameter.

@@

The syntax of the inheritance judgement is very close to how Family is defined in the surface syntax -- a Family with no parent is an inheritance with empty input; an extended Family is exactly an inheritance -- because the parent of an extended Family can be overridden thus inheritance should only define upon the ``interface'' instead of a concrete family. However, since our current formulation doesn't support further-binding (i.e. we cannot do inheritance inside a family), our formulation still have some distance to the real family polymorphism.

"inhid" and "inhext" are simple as expected -- the former corresponds to empty inheritance, and the latter gives the programmer the ability to add new field.

@@

Override "inhov" is even more special because in mundane OO programming language, we usually require that the overriding term has the same type/interface as the overriden one.  Here we don't since the very reason we want the same type/interface is that we hope other inherited fields can use this overridden/late-binding field without breaking abstraction, and this reason is already managed by the assumption of "inhinh", the upcasting proof $\uparrow^s$. Thus "inhov" is quite like "inhext", where overriding is just throwing away one parent field. This difference also show one of the distinct feature of our vanilla family inheritance compared to the mainstream OO inheritance -- family inheritance don't need to introduce the concept of ``subclassing'' between families and thus no bounded polymorphism yet\footnote{Of course once trait and interface comes in, we can have bounded polymorphism back}, and code reuse is mainly achieved by late binding. Without ``subclassing'', there is no way of confusing inheritance and subtyping.

>>>>

Firstly, there are four standard judgement forms in MLTT,
well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
$\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. Our presentation
follows the style of~\citet{kaposi2019gluing}, but for readability we
omit universe levels attached to judgements.
After these,
we will have four new kinds of judgement to model our linkage.

@@

Due to our intrinsic setting, \textbf{we can and will omit a lot of presumptions of the typing rules}, for example we don't spell out that the rule "Type Universe" requires a well-typed context, even though it did implicitly---because it is not possible to have "not-well-formed" context in our setting.

@@

\judgebox{\goodWSig{\Gamma}{\sigma}{n}}
$$
\Rule[name=Emp-WSig]
{}
{\goodWSig{\Gamma}{w\cdot}{0}}
\quad
\Rule[name=WSig-Add]
{\goodWSig{\Gamma}{s}{n}
  \quad \goodType{\Gamma}{A}{i}
  \quad \goodType{\Gamma, A}{B}{i}}
{\goodWSig{\Gamma}{w^+ \  s \  A \  B}{n+1}}
\quad
\Rule[name=Ind-Univ]
{\goodWSig{\Gamma}{\sigma}{n}}
{\goodType{\Gamma}{\bW \sigma}{i}}
$$
%\YZ{Is there an introduction rule for terms of type 'bW sigma'?}
%\EDJreply{Yes there is. I omit it because it is a singleton type, please see my newly added explanation below.}

@@

Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A \vdash B(x)$ decides a W-type. To simulate extensible inductive type, we have to equip our inductive type with multiple constructors, and for simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ : (x : Aᵢ) → (Bᵢ x → W) → W}. Thus we use a list of pairs of such $A_i, B_i$ as the signature for one inductive type, and given an inductive type signature, we use \ruleref{WSig Proj} to extract and corresponding $A_i, B_i$.

Of course we need to put an eye on the substitution law for each piece of the syntax. Here $\bW \sigma$ is a large singleton type~\cite{stone2000}, and it only has one ``element'', the inductive type itself, inside. We formulate in this cumbersome way because we want to have inductive type to be a field element and exposing the constructors at the type level for later \textit{pattern-matching coverage checking} (i.e. if an inductive type as field element has type $\cU$, then no information about the inductive type is exposed).\YZ{What is the distinction of 'W T' vs 'El (W T)'?}\EDJreply{(W T) is a term inside the type universe, El (W T) is really a type. They are basically the same thing but (W T) cannot be at the type position in a judgement, i.e. k |- q : (W T) is not allowed, we only have k |- q : El (W T). This is kind of standard notation (at least in both Sterling and Kaposi) so I will avoid explaining in the main text if you think ok?}\YZreply{I think explanation is warranted; readers unfamiliar with that line of work may find it confusing. BTW, is the elimination rule for "El (W T)" omitted?}\EDJreply{Please check the following paragraph for the explanation for El, and also Univ-1 is the elimination rule for "El T" in general. I just didn't write out the two equations c (El T) = T, El (c T) = T}. We follow the style of \textit{type universe à la Tarski} and distinguish types from their \textit{codes or names}, and thus we use $El\ T$ as the type given type name $T : \cU$---in these cases, $T$ is not allowed to locate at type position (i.e. at the right-most position of a term judgement) but $El \ T$ is.

@@

The signature (for the linkage) and the linkage rules are quite similar the signature for inductive type.  They are both constructed from the empty stuff, and respect substitution lemma for sure. We have \ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each componenet due to harmony~\cite{pfenning2009lecture} thus we have local soundness and local completeness ($\beta,\eta$ rules). We omit them and the substitution laws for either field addition and projection. For $\mu^+,\nu^+$ rules, we will omit $\{f\}$ if possible.\footnote{This omission can be supported by Agda using Implicit Arguments}

@@

In our meta-theory, we only formulate non-dependent eliminator, i.e. the
\textit{recursor}. Each recursor is constructed by a linkage containing
functions/handlers/pattern-matching cases dealing with each
constructor. Thus code reuse for recursor is delegated to linkage
inheritance----to reuse one particular pattern-matching case,
we just inherit the function/handler. This greatly simplifies the
metatheory development.

@@

Now, one missing piece for linkage is the sealing judgement, and both of the $\mu^+, \nu^+$ rules relies on that---then what is this sealing judgement for?

@@

As we said earlier, $\goodSeal{\Gamma}{\_}{\sigma}{A}$ is actually just
a short hand of term judgement $\goodTerm{\Gamma , \cC
\sigma}{\_}{A[\pi_1]}$, a special form of \textit{local morphism}~\cite{abbott2003category}. The substitution law should be understood as a
lemma, and it is easily derivable using $\gamma^\uparrow$.

The motivation for sealing judgement is acting like a syntactic sugar helpful for inheritance---due to the presence of inductive type in the context, not everything is inheritable. For example, $ X_1 : \bW \sigma_1 \vdash t : T$ in the parent family will generally not be able to be inherited to the children once we override the inductive type. In other words, there is no way ``transplanting'' $t$ into the context $X_2 : \bW \sigma_2$. This makes all the fields after this inductive type non-inheritable---which shouldn't be the case because a lot of fields might not be directly relying on this concrete $\bW \sigma_1$.

Sealing is used to exploit this fact---with sealing, we can add well-typed terms $X_1 : \cU \vdash t : T$, into the parent family, in the context of $ X_1 : \bW \sigma_1$. Later when we want to inherit $t$, because $t$ only consider $X_1$ as an arbitrary type, inheritance is plausible.

Of course, any fields using "Wsup" and "Wrec" directly relies on the $\bW(\sigma)$ (can be spotted from \ruleref{Ind-Term} and \ruleref{Rec-Constr}) and thus non-inheritable---in these cases, "Wsup" can be understood as concrete constructors and "Wrec" are \textit{exhaustiveness checking}. These fields together with inductive types are the one sealing responsible for hiding. After compilation of linkage type into existential type, we can easily hide the \textit{concrete} definition of the inductive type while exposing all the fields using "Wsup" and "Wrec" like an abstract interface. Apparently, future fields relies on this abstract interface will have no problem on inheritance. \YZ{Does coverage checking mean exhaustiveness checking?}\EDJreply{Yes. The one make sure every case is handled in pattern-matching. Do you want me to replace the terminology using exhaustiveness checking? }\YZreply{Yes, I think exhaustiveness is standard parlance.}
Of course, since "Wsup" and "Wrec" directly rely on the concrete inductive definition, we will have to override these fields whenever inductive type is overridden. For "Wsup" we only need to override each constructor and for "Wrec" it is just doing exhaustiveness checking inside each children family. The recursor handlers however, can be safely inherited.
\YZ{I suppose the terminology 'sealing' is a reference to ML modules? In
ML, sealing has the connotation of hiding things behind an existential
type. So using the same word may cause the reader to have false
presumptions about what is going on here.}\EDJreply{You are right. Our "Sealing" is not hiding a particular field into an existential type, but our seal is make several fields together into an "existential type". Basically it hides the concrete definition of inductive type so that future fields doesn't have to be defined upon that.}\EDJreply{I will resolve this comment once I add the throughout example and you find it clear---because I haven't changed any text}

@@

With the definition of sealing judgement in mind, we can see
\ruleref{Sig Add} and \ruleref{Lnkg Add} work exactly as the
implementation. In the implementation, we will have "self__" all around
when defining new fields in our plugin, and this "self__" is exactly the
$A$ in the context of the two fields---especially when we don't do any
sealing, this $A = \cC \sigma$, which is exactly how the implementation
exposes "self__" as a module parameter.

@@

The syntax of the inheritance judgement is very close to how Family is defined in the surface syntax---a Family with no parent is an inheritance with empty input; an extended Family is exactly an inheritance---because the parent of an extended Family can be overridden thus inheritance should only define upon the ``interface'' instead of a concrete family. However, since our current formulation doesn't support further-binding (i.e. we cannot do inheritance inside a family), our formulation still have some distance to the real family polymorphism.

"inhid" and "inhext" are simple as expected---the former corresponds to empty inheritance, and the latter gives the programmer the ability to add new field.

@@

Override "inhov" is even more special because in mundane OO programming language, we usually require that the overriding term has the same type/interface as the overriden one.  Here we don't since the very reason we want the same type/interface is that we hope other inherited fields can use this overridden/late-binding field without breaking abstraction, and this reason is already managed by the assumption of "inhinh", the upcasting proof $\uparrow^s$. Thus "inhov" is quite like "inhext", where overriding is just throwing away one parent field. This difference also show one of the distinct feature of our vanilla family inheritance compared to the mainstream OO inheritance---family inheritance don't need to introduce the concept of ``subclassing'' between families and thus no bounded polymorphism yet\footnote{Of course once trait and interface comes in, we can have bounded polymorphism back}, and code reuse is mainly achieved by late binding. Without ``subclassing'', there is no way of confusing inheritance and subtyping.
